---
title: "Introduction to stathelper"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to stathelper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stathelper)
```

# stathelper: Statistical Helper Functions

The `stathelper` package provides simple utility functions for common statistical operations. This vignette demonstrates how to use each function in the package.

## Calculating Z-Scores

Z-scores (also called standard scores) tell you how many standard deviations away from the mean a data point is. The `z_score()` function calculates z-scores for a vector of values.

```{r}
# Create a sample dataset
data <- c(23, 19, 27, 31, 22, 25, 18, 29)

# Calculate z-scores
z_scores <- z_score(data)
z_scores

# Verify the mean and sd of the z-scores
mean(z_scores)
sd(z_scores)
```

You can also provide your own mean and standard deviation:

```{r}
# Calculate z-scores with specified parameters
z_score(c(85, 90, 95, 100), mean = 90, sd = 5)
```

## Calculating Confidence Intervals

The `confidence_interval()` function calculates the confidence interval for the mean of a data set:

```{r}
# Get a confidence interval for our sample data
ci <- confidence_interval(data)
ci

# The default confidence level is 0.95, but we can change it:
confidence_interval(data, conf.level = 0.99)
```

## Detecting Outliers

The `detect_outliers()` function uses the IQR method to identify outliers in your data:

```{r}
# Create data with outliers
outlier_data <- c(23, 19, 27, 31, 22, 25, 18, 29, 95)

# Detect outliers
outliers <- detect_outliers(outlier_data)
outliers

# Get the outlier values
outlier_data[outliers]
```

We can adjust the sensitivity of the outlier detection by changing the multiplier:

```{r}
# More strict outlier detection
detect_outliers(outlier_data, multiplier = 1)

# More lenient outlier detection
detect_outliers(outlier_data, multiplier = 3)
```
